Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})

### 2014-05-19 manuscript.tex
edits to discussion

Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.  Drift may have played a role in the San Lorenzo \emph{parviglumis} population. Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele at \emph{tb1} could have risen to high frequency entirely by chance. This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} allele in multiple populations in the Jalisco group.  Moreover, available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.

\jri{here we need a few sentences on selection.  the 5' UTR has much more negative D than the upstream. do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the \emph{Hopscotch} insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..

>>>>

Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.  Drift may have played a role in the San Lorenzo \emph{parviglumis} population. Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele at \emph{tb1} could have risen to high frequency entirely by chance. This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} allele in multiple populations in the Jalisco group.  Moreover, available information on diversity and population structure among additional Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.  Values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.  Whereas the genome-wide average of Tajima's D in genic regions of \emph{parviglumis} is 0.45 \citep{Hufford2012b}, the statistic is quite negative in the 5' UTR of \emph{tb1} (Table \ref{Table2Diversity}). This result is consistent with repeated selective sweeps near \emph{tb1} and an important ecological role for the gene in \emph{parviglumis}.

\jri{do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
